---
title: "Pset 4"
author: "Diego Martinez"
date: "2/24/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(gt)
library(readr)
library(lubridate)
library(ggplot2)
library(forcats)
library(fivethirtyeight)

#loading in data, not filtering out 3rd party candidates becuase they will be important in later calculations 
poll_data <- read_csv("ps_4_elections-poll-nc09-3.csv",
col_types =  cols(
                   .default = col_character(),
                   turnout_scale = col_double(),
                   turnout_score = col_double(),
                   w_LV = col_double(),
                   w_RV = col_double(),
                   final_weight = col_double(),
                   timestamp = col_datetime(format = ""))) 


```


```{r, echo=FALSE} 
# grouping by responses then turning the rows into columns to more easily access the values using the pull function
x<-poll_data %>% 
  group_by(response)%>%
  count()%>%
  spread(key = response, value = n)


ML1<- x %>%
  pull(Dem)

ML2<-x$Rep - x$Und

ML3<-poll_data %>% 
  filter(gender != gender_combined) %>% tally()

#I have gotten two answers for this question depending if I filter out 3rd parties. Right now I did not filter out third parties at got 28; however when I do filter them out the answer is 27.
ML4 <-poll_data %>% 
  filter(race_eth == "White" & file_race_black != "White")%>% 
  tally()

#Dem response came before repubiclans first response, I changed the ML to say "Rep came 11 minutes after (used to say before) the first response for Dem"
ML5<-poll_data%>% 
  filter(response %in% c("Dem", "Rep"))%>% 
  group_by(response)%>% 
  select(response, timestamp) %>% 
  summarize(first = min(timestamp)) %>% 
  summarize(diff = round(first[2]- first[1]))





```
## Question 1

There were `r ML1` respondents who supported the Democratic candidate.

There were `r ML2` more respondents who favored the Republican candidate than who were Undecided.

There are two gender variables (gender and gender_combined). There are `r ML3` individuals for whom these variables have different values.

There are `r ML4` respondents listed as “White” under race_eth who are not listed as “White” under file_race_black.

The first response of Rep came `r ML5` minutes (rounded to the nearest minute) after the first response of Dem.


## Question 2

```{r, echo= FALSE}

z <-
  #filtering to remove the don't know/refused because they change the percentages and create a new row of data in the end table. 
  poll_data %>%  filter(!is.na(response), race_eth != "[DO NOT READ] Don't know/Refused") %>%
  # it is not intuitive to r which should be displayed first because they are factor variables, thus I changed the order to match the example table 
  mutate(race_eth = fct_relevel(race_eth, c(
  "White", "Black", "Hispanic", "Asian", "Other"
  ))) %>%
  #taking out columns unnecessary for this analysis 
  select(race_eth, response, final_weight) %>%
  # It was important to group by race_eth first before responses to get correct counts and percentages. 
  group_by(race_eth, response) %>%
  # not all the respondents are weighted the same based on the various answers to the poll, 
  summarize(total = sum(final_weight)) %>%
  #turns the column from the group_by of responses into a row of responses with each column being a different response to create a percentage/ probability table where one variable is racce_eth and the other is response.
  spread(key = response, value = total, fill = 0) %>%
  #need to add  all responses together for a grand total to begin creating percentages
  mutate(y = Dem + Rep + Und + `3`) %>%
  mutate(Dem = (Dem) / y) %>%
  mutate(Rep = (Rep) / y) %>%
  mutate(Und = (Und) / y) %>%
  #leaving out the 3rd party responeses because they are not included in the example
  select(race_eth, Dem, Rep, Und) %>%
  #turns values that are 0 into NA because there are no Asians that are undecided
  na_if(0) %>%
  #avoiding any potential errors from the groupings I made before
  ungroup() %>%
  gt() %>%
  tab_header("Demographics of North Carolina’s 9th Congressional District") %>%
  #crediting the source of the data
  tab_source_note("New York Times Upshot/Siena College 2018 live polls") %>%
  cols_label(
  race_eth = "Ethnicity",
  Dem = "DEM.",
  Rep = "REP.",
  Und = "UND."
  ) %>%
  # rounds and cleans up the percentages to make the table nicer and resemble example table more.
  fmt_percent(columns = vars(Dem, Rep, Und), decimals = 0)%>%
  # changes the 0 in Und Asians box to a -
  fmt_missing(columns = vars(Und), rows =4) 
  
  z

  
```

##Question 3

```{r, echo=FALSE, fig.width = 7.2}

Q3 <- poll_data %>%
  filter(!is.na(response), educ != "[DO NOT READ] Refused") %>%
  #the order is not obvious to r from Grade school being lowest to Graduate or Professional Degree being the hgihest
  mutate(
  educ = fct_relevel(
  educ,
  "Grade school",
  "High school",
  "Some college or trade school",
  "Bachelors' degree",
  "Graduate or Professional Degree"
  )
  ) %>%
  #keeping together all those that have same education level
  group_by(educ) %>%
  # the axis are originally flipped because if not geom_violin will not work.
  ggplot(aes(y = final_weight, x = educ)) + 
  geom_violin() + 
  # changing the width and alpha makes it more resemble the example graph because the data points were more spread out and a darker transparency.
  geom_jitter(alpha = 0.4, width = .2) +
  #puts levels of education on the y axis. 
  coord_flip() +
  labs(
  x = NULL,
  y = "Weight Given to Respondent in Calculating Poll Results",
  title = "More Educated Matter Less in North Carolina 9th",
  subtitle = "Poll gives more weight to people who are less likely to participate in polls",
  caption = "New York Times Upshot/Siena College 2018 live polls"
  )
  
  Q3


```



##Question 4
```{r, echo=FALSE, fig.width=7.2}
poll_data %>%
  #cleaning data, I only want to compare those that are decided, thus I only keep Democratic and Republican responses 
  filter(!is.na(response),file_race != "[DO NOT READ] Don't know/Refused",response %in% c("Dem", "Rep"),ager != "[DO NOT READ] Refused") %>%
  #the order of the groupings matters, the way I have grouped them is to show how each gender for each age range.
  group_by(gender, ager , response) %>%
  summarise(N = n()) %>%
  #turning into a percentage to be able to graph categorical data
  mutate(freq = N / sum(N) , pct = round((freq * 100) , 0)) %>%
  ggplot(aes(x = ager, y = pct, fill = gender)) + geom_col(position = "dodge2") +
  #facet wrap to compare the parties
  facet_wrap( ~ response) + labs(
  y = "Percentage of Vote by Gender per Age Range",
  title = "Gender's Vote Differently in NC 9th",
  subtitle = "Dems Win Every Female Age Range While Reps Win Every Male Age Range",
  x = "Age Range",
  caption = "New York Times Upshot/Siena College 2018 live polls"
  )
  



```


##Colleagues
Baeu M, Chris O, Chris M, Debora G

